Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2021-04-17 manuscript/manuscript.Rmd
Minor metadata updates
Super minor things

date: "October 2, 2020"

@@

  The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income to INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors, and trust has shown to be an important component of support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals’ preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.

@@

word-count: "7,562"

@@

As a result of these laws, public or official foreign aid channeled through international NGOs (INGOs) has decreased considerably in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. However, philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018]. Given the reduction of foreign aid following restricted legal environments abroad, INGOs may need to devote more resources to fundraise from individual donors.

>>>>

date: "April 17, 2021"

@@

  The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income to INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors, and trust has shown to be an important component of support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals' preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.

@@

word-count: "7,986"

@@

As a result of these laws, public or official foreign aid channeled through international NGOs (INGOs) has decreased in repressive countries [@Brechenmacher:2017; @ChaudhryHeiss:2018; @DupuyPrakash:2018]. However, philanthropy from private donors and foundations is not as adversely affected [@McGill:2018]. Many foundations have continued channeling funds to countries with hostile legal environments [@NeedlesWolcheckSato:2018]. Given the reduction of foreign aid following restricted legal environments abroad, INGOs may need to devote more resources to fundraise from individual donors.
